---
title       : dplyr Review
description : Create a new column in a data frame, modify an existing column, and sort by one or more columns using the dplyr package.  Review the main ideas of the Five Main Verbs of dplyr - filter, summarize, group_by, mutate, and arrange (in development).

--- type:NormalExercise lang:r xp:100 skills:1,4 key:62a044eae1
## Create two new variables

Recall the `bechdel` data set in the `fivethirtyeight` package used in the 2014 article from FiveThirtyEight.com entitled "The Dollar-And-Cents Case Against Hollywood’s Exclusion of Women" [here](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/).  

*** =instructions
- First create a modified data frame that includes only years 2000 to 2013, inclusive.  Provide the name `bechdel00_13` for this smaller data frame.  Note that the `between()` function in `dplyr` may be helpful.  You should check that you have correctly created `bechdel00_13` in the R Console.
- Now add a column to `bechdel00_13` that calculates the domestic return on investment for all 2000 to 2013 movies based on 2013 dollars.  This is calculated as `roi_dom = domgross_2013 / budget_2013` and gives how much money the movie earned for each dollar spent.


*** =hint
- `between` expects three arguments:  the name of the variable, the beginning entry, and the ending entry.
- Remember to add a variable to a data frame you need to assign the name back to include the change.  Something like `flights <- flights %>% mutate(gain = arr_delay - dep_delay)`.

*** =pre_exercise_code
```{r}

```

*** =sample_code
```{r}

```

*** =solution
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel00_13 <- bechdel %>%
    filter(between(year, 2000, 2013)) %>%
    mutate(roi_dom = domgross_2013 / budget_2013)
```

*** =sct
```{r}
test_library_function("fivethirtyeight")
test_library_function("dplyr")
test_object("bechdel00_13")
success_msg("Nicely done!")
test_error()
```