---
title       : dplyr Review
description : Create a new column in a data frame, modify an existing column, and sort by one or more columns using the dplyr package.  Review the main ideas of the Five Main Verbs of dplyr - filter, summarize, group_by, mutate, and arrange (in development).

--- type:NormalExercise lang:r xp:100 skills:1,4 key:62a044eae1
## Create a new variable

Recall the `bechdel` data set in the `fivethirtyeight` package used in the 2014 article from FiveThirtyEight.com entitled "The Dollar-And-Cents Case Against Hollywood’s Exclusion of Women" [here](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/).  

*** =instructions
- First create a modified data frame that includes only years 2000 to 2013, inclusive.  Provide the name `bechdel00_13` for this smaller data frame.  You should check that you have correctly created `bechdel_00_13` in the R Console.
- Now add a column to `bechdel_00_13` that calculates the domestic return on investment for all 2000 to 2013 movies based on 2013 dollars.  This is calculated as `roi_dom = domgross_2013 / budget` and gives how much money the movie earned for each dollar spent.


*** =hint
- You may want to use the `between` function in `dplyr` for your `filter` here, which expects three arguments:  the name of the variable, the beginning entry, and the ending entry.
- Remember to add a variable to a data frame you need to assign the name back to include the change.  Something like `flights <- flights %>% mutate(gain = arr_delay - dep_delay)`.

*** =pre_exercise_code
```{r}

```

*** =sample_code
```{r}

```

*** =solution
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel_00_13 <- bechdel %>%
    filter(between(year, 2000, 2013)) %>%
    mutate(roi_dom = domgross_2013 / budget)
```

*** =sct
```{r}
test_library_function("fivethirtyeight")
test_library_function("dplyr")
test_object("bechdel_00_13")
success_msg("Nicely done!")
test_error()
```

--- type:NormalExercise lang:r xp:100 skills:1,4 key:4de37aa411
## Modify an existing variable

We will continue using the `bechdel_00_13` data set you created in the last exercise.  It has been loaded for you here.

We used the `budget` variable in creating the `roi_dom` variable in the last exercise.  What we should have used is the `budget_2013` variable so that both the domestic gross and the budget were on the same 2013 level.  We can adjust `roi_dom` to have the correct value by multiplying by `budget_2013` and dividing by `budget`.

*** =instructions
- Modify the `roi_dom` variable in `bechdel_00_13` to include the correct entries for `roi_dom` as discussed above.


*** =hint
- Remember to use the `mutate()` function to modify the contents of a variable in a data frame.

*** =pre_exercise_code
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel_00_13 <- bechdel %>%
    filter(between(year, 2000, 2013)) %>%
    mutate(roi_dom = domgross_2013 / budget)
```

*** =sample_code
```{r}
library(dplyr)

```

*** =solution
```{r}
library(dplyr)
bechdel_00_13 <- bechdel_00_13 %>%
    mutate(roi_dom = budget_2013 / budget)
```

*** =sct
```{r}
test_library_function("dplyr")
test_object("bechdel_00_13")
success_msg("Nicely done!")
test_error()
```