---
title       : Applying R Basics
description : Exploring the basics of R on a data set based on the fivethirtyeight.com "Most Police Don’t Live In The Cities They Serve" article.


--- type:NormalExercise lang:r xp:100 skills:1 key:c306f947db
## Investigate the properties of a data frame

Read over the 2014 article from FiveThirtyEight.com entitled "Most Police Don’t Live In The Cities They Serve" [here](http://fivethirtyeight.com/datalab/most-police-dont-live-in-the-cities-they-serve/).   Police and their relationships with citizens has been a controversial topic for many years and has come to the forefront of American culture since the events of Ferguson, Missouri mentioned in this article.

The dataset corresponding to this article from the `fivethirtyeight` package has been loaded as a data frame with name `police_locals` and is available in the workspace.  You can view the `police_locals` data frame as an interactive table [here](https://ismayc.github.io/Effective-Data-Storytelling-using-the-tidyverse/police_locals.html).

To work with functions and/or data stored in an R package, we need to first load the package using the `library` function.  This is done for you with both the `tibble` package and the `fivethirtyeight` package.  The `data` function loads the specified data set (passed as an argument to `data`) into the environment.

Remember to use the R Console as your sandbox for checking that your R commands do what you suspect them to do.  When you are confident with your answer, enter it into **script.R** and press Submit Answer.  You can **Take Hint** but you should rarely be using this feature as you'll find you will better learn the material if you struggle with the problems some.

*** =instructions
- Get the names of the columns/variables in `police_locals` using the `names` function.
- Check out the properties of `police_locals` using the `glimpse` function in the `tibble` package.  (This function acts much like the `str` function, but has some nicer features.)

*** =hint
- Use `names()`.  Note that what goes inside the `()` is the name of the data frame you'd like to examine the properties of.
- Use `glimpse()`.  Note that what goes inside the `()` is the name of the data frame you'd like to examine the properties of.
- Note that you should NOT put the name of the data frame in quotation marks.  Use `?names` and/or `?tibble::glimpse` for more information.

*** =pre_exercise_code
```{r}
library(tibble)
library(fivethirtyeight)
data(police_locals)
#library(readr)
#police_locals <- read_csv("https://raw.githubusercontent.com/ismayc/Effective-Data-Storytelling-using-the-tidyverse/master/datasets/police_locals.csv")
```

*** =sample_code
```{r}
library(tibble)
library(fivethirtyeight)
data(police_locals)
# police_locals is available in your workspace
# Identify the names of the columns in police_locals

# Check out the properties of police_locals

```

*** =solution
```{r}
library(tibble)
library(fivethirtyeight)
data(police_locals)
# police_locals is available in your workspace
# Identify the names of the columns in police_locals
names(police_locals)
# Check out the properties of police_locals
glimpse(police_locals)
```

*** =sct
```{r}
# SCT written with testwhat: https://github.com/datacamp/testwhat/wiki
test_function("names", args = "x",
              not_called_msg = "You didn't call `names()`!",
              incorrect_msg = "You didn't call `names(x = ...)` with the correct argument, `x`.")

test_function("glimpse", args = "x",
              not_called_msg = "You didn't call `glimpse()`!",
              incorrect_msg = "You didn't call `glimpse(x = ...)` with the correct argument, `x`.")

test_error()

success_msg("Good work! <br><br> THINK ABOUT IT:  Why are the variables in `police_locals` the specific `class` they are coded as here (`chr`, `int`, `dbl`)?")
```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:792856c06f
## Identify observational unit

What is the observational unit in this `police_locals` data frame?  You can view the `police_locals` data frame as an interactive table [here](https://ismayc.github.io/Effective-Data-Storytelling-using-the-tidyverse/police_locals.html).

*** =instructions
- Percentage of officers
- City
- Police force
- Race

*** =hint
What is the thing being measured along each row?  Remember that variables (unless they are identification variables) cannot be the same
as the *observational unit*.

*** =pre_exercise_code
```{r}
# You can also prepare your dataset in a specific way in the pre exercise code
library(fivethirtyeight)
data(police_locals)
library(tibble)
```

*** =sct
```{r}
# SCT written with testwhat: https://github.com/datacamp/testwhat/wiki

msg_bad <- "That is not correct!"
msg_success <- "Exactly! The observational unit here is city.  Percentage of officers, police force, and race are all variables/characteristics about the city."
test_mc(correct = 2, feedback_msgs = c(msg_bad, msg_success, msg_bad, msg_bad))
```

--- type:NormalExercise lang:r xp:100 skills:1 key:a713e275ff
## Create a vector

One of the cool features of R is that works quickly and easily with vectors.  Test your knowledge from the Introduction to R DataCamp course
by determining the vector specified below.

*** =instructions

- Create a new vector called `all_not` that calculates the percentage of the total police force that **DO NOT** live in the city.  Note that `all_not` should not be included in the `police_locals` data frame but instead live as a vector by itself.
- You can get information on what each of the different variables in `police_locals` represents by entering `?police_locals` in the **R Console** or by going [here](https://rdrr.io/cran/fivethirtyeight/man/police_locals.html).
- NOTE: The best way to extract a vector from a data frame is to use `$`.


*** =hint
- Remember that you can add or subtract a value from all elements of a vector using something like `vec2 <- 2 + df$vec1` where `vec1` is a vector
in the `df` data frame.
- The total police force is stored in the vector `all` in `police_locals`.
- Use the assignment operator `<-` to give a name to an object (such as a vector).


*** =pre_exercise_code
```{r}
library(tibble)
library(fivethirtyeight)
data(police_locals)
```

*** =solution
```{r}
all_not <- 1 - police_locals$all
```

*** =sct
```{r}
test_object("all_not")
test_error()
```



--- type:NormalExercise lang:r xp:50 skills:1 key:d4dc193a51
## Subset and index a vector

Use `[ ]` to extract specific information about a variable in the `police_locals` data frame.

*** =instructions
- Extract the first, seventh, sixty-sixth, and seventy-third city names from the data set in **one** R command.
- Assign `sub_cities` as the name to this object

*** =hint
- Remember the `c` function is used to provide a vector of values separated by columns.
- Note that it is only asking for entries in the `city` variable, not any other variables.
- Recall to extract a vector from a data frame, we use `$`.

*** =pre_exercise_code
```{r}
library(tibble)
library(fivethirtyeight)
data(police_locals)
```

*** =solution
```{r}
sub_cities <- police_locals$city[c(1, 7, 66, 73)]
```


*** =sct
```{r}
test_object("sub_cities")
test_error()
```


--- type:NormalExercise lang:r xp:100 skills:1 key:bd0f6596e5
## Subset and index a vector again.

In R, an index refers to the specific entry in an object.  As an example, in a vector called `my_vec` defined as
`my_vec <- c(200, 243, 78)`, index 1 corresponds to 200, index 2 corresponds to 243, and index 3 corresponds to 78.

*** =instructions
Extract the police size forces corresponding to indices 14 to 39 in `police_locals` and assign the `pol_force` name to this subset.

*** =hint
- Remember police size forces are giving in the `police_size_force` vector.
- The `:` operator can be used to specify a range of values.  You can look for help on this by entering `?` `` `:` `` in the R Console.
- You don't have to use `c` with `:`.  For example, `my_vec2 <- my_vec[c(1:2)]` sends the first two indexed values of `my_vec` to an object with the name `my_vec2`, but `my_vec2 <- my_vec[1:2]` also works and is preferred.

*** =pre_exercise_code
```{r}
library(tibble)
library(fivethirtyeight)
data(police_locals)
```

*** =solution
```{r}
pol_force <- police_locals$force_size[14:39]
```


*** =sct
```{r}
test_object("pol_force")
test_error()
```