---
title       : ggplot2 Review
description : Identifying and creating the appropriate plot based on different types of variables

--- type:NormalExercise lang:r xp:100 skills:1 key:f6d3394a54
## Fixing an inappropriate plot

When one doesn't properly use the correct plot based on the type of variables, some pretty strange plots can be produced.  These plots are hard to interpret and often extremely ineffective at telling a story about the data.  This exercise and the ones that follow will have you examine why a plot is incorrect and/or fix incorrect code to make a more appropriate plot.

Read over the 2014 article from FiveThirtyEight.com entitled  "The Economic Guide To Picking A College Major" [here](http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/).  You can find the data corresponding to the story as `college_all_ages` in the `fivethirtyeight` R package.  We are interested in looking at the distribution of the `unemployment_rate` variable in this `college_all_ages` data frame.

*** =instructions
- Fix the code given to produce an appropriate plot looking at how `unemployment_rate` varies in the `college_all_ages` data frame.
- Make sure to also use the `color = "white"` argument to better
clarify your plot.

*** =hint
- What type of variable is `unemployment_rate`?
- The Coggle mind map may be of assistance [here](https://coggle.it/diagram/V_G2gzukTDoQ-aZt-).

*** =pre_exercise_code
```{r}
library(fivethirtyeight)
library(ggplot2)
ggplot(data = college_all_ages, mapping = aes(x = unemployment_rate)) +
  geom_bar()
```

*** =sample_code
```{r}
library(fivethirtyeight)
library(ggplot2)
ggplot(data = college_all_ages, mapping = aes(x = unemployment_rate)) +
  geom_bar()
```

*** =solution
```{r}
library(fivethirtyeight)
library(ggplot2)
ggplot(data = college_all_ages, mapping = aes(x = unemployment_rate)) +
  geom_histogram()
```

*** =sct
```{r}
test_library_function("ggplot2")
test_ggplot(check_data = TRUE, check_aes = TRUE, check_geom = TRUE,
            check_geom_params = TRUE)
success_msg("Nicely done! <br><br> THINK ABOUT IT:  Why is a barplot inappropriate when looking at the distribution of a numerical variable?")
test_error()
```