---
title       : dplyr Review
description : Create a new column in a data frame, modify an existing column, and sort by one or more columns using the dplyr package.  Review the main ideas of the Five Main Verbs of dplyr - filter, summarize, group_by, mutate, and arrange.

--- type:NormalExercise lang:r xp:100 skills:1,4 key:762c02a329
## Modify an existing variable

Recall the `bechdel` data set in the `fivethirtyeight` package used in the 2014 article from FiveThirtyEight.com entitled "The Dollar-And-Cents Case Against Hollywood’s Exclusion of Women" [here](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/).  As it currently stands in the `fivethirtyeight` package, the `clean_test` variable is just a character vector.  The entries do have a logical order though and should, thus, be converted to a `factor` variable using the `levels` argument to set them as  

```
levels = c("nowomen", "notalk", "men", "dubious", "ok")
```

*** =instructions
- Use the `mutate` function to change `clean_test` to be defined as a `factor` with `levels` defined above.
- Assign this modified data frame to have the same name of `bechdel`.


*** =hint
- Remember to check the documentation for `factor` via `?factor` if you need assistance.



*** =pre_exercise_code
```{r}

```


*** =sample_code
```{r}

```

*** =solution
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel <- bechdel %>%
    mutate(clean_test = factor(clean_test, 
                levels = c("nowomen", "notalk", "men", "dubious", "ok")))
```

*** =sct
```{r}
test_library_function("fivethirtyeight")
test_library_function("dplyr")
test_object("bechdel")
success_msg("Nicely done!")
test_error()
```



--- type:NormalExercise lang:r xp:100 skills:1,4 key:62a044eae1
## Create a new variable

You will use the modified `bechdel` data frame you created in the last exercise here.  It has been loaded for you.

*** =instructions
- First create a modified data frame that includes only years 2000 to 2013, inclusive, from the `bechdel` data frame.  Provide the name `bechdel_00_13` for this smaller data frame.  You should check that you have correctly created `bechdel_00_13` in the R Console.
- Now add a column to `bechdel_00_13` that calculates the domestic return on investment for all 2000 to 2013 movies based on 2013 dollars.  This is calculated as `roi_dom = domgross_2013 / budget_2013` and gives how much money each movie earned for every dollar spent.


*** =hint
- You may want to use the `between` function in `dplyr` for your `filter` here, which expects three arguments:  the name of the variable, the beginning entry, and the ending entry.
- Remember to add a variable to a data frame you need to assign the name back to include the change.  Something like `flights <- flights %>% mutate(gain = arr_delay - dep_delay)`.

*** =pre_exercise_code
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel <- bechdel %>%
    mutate(clean_test = factor(clean_test, 
                levels = c("nowomen", "notalk", "men", "dubious", "ok")))
```

*** =sample_code
```{r}

```

*** =solution
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel_00_13 <- bechdel %>%
    filter(between(year, 2000, 2013)) %>%
    mutate(roi_dom = domgross_2013 / budget_2013)
```

*** =sct
```{r}
test_library_function("fivethirtyeight")
test_library_function("dplyr")
test_object("bechdel_00_13")
success_msg("Nicely done!")
test_error()
```



--- type:NormalExercise lang:r xp:100 skills:1,4 key:84af7c1ecd
## Aggregate based on values of another variable

We will continue using the `bechdel_00_13` data set you created in the last exercise.  It has been loaded for you here.

Let's now see how movies that passed the Bechdel test faired in terms of return on investment to those that didn't pass.  Recall that the `binary` variable stored whether or not a movie passed the Bechdel test.

*** =instructions
- Determine the mean and median values for `roi_dom` for those that passed the Bechdel test and those that failed.
- Call these new variables in the aggregated data frame `mean_roi` and `median_roi` respectively.
- Give the name of `roi_summ` to the aggregated data frame and print it to the Console.

*** =hint
- You should be using `group_by` and `summarize` here.

*** =pre_exercise_code
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel <- bechdel %>%
    mutate(clean_test = factor(clean_test, 
                levels = c("nowomen", "notalk", "men", "dubious", "ok")))
bechdel_00_13 <- bechdel %>%
    filter(between(year, 2000, 2013)) %>%
    mutate(roi_dom = domgross_2013 / budget)
bechdel_00_13 <- bechdel_00_13 %>%
    mutate(roi_dom = budget_2013 / budget)    
```

*** =sample_code
```{r}

```

*** =solution
```{r}
library(dplyr)
roi_summ <- bechdel_00_13 %>%
    group_by(binary) %>%
    summarize(mean_roi = mean(roi_dom),
              median_roi = median(roi_dom))
roi_summ
```

*** =sct
```{r}
test_library_function("dplyr")
test_object("roi_summ")
success_msg("Nicely done! <br> <br> THINK ABOUT IT:  What can be said about the claim that some movie executives make that movies that
have women playing strong roles in movies don't make as much return on investment than those that don't based on this data?")
test_error()
```

--- type:NormalExercise lang:r xp:100 skills:1,4 key:5e09c823f9
## Aggregate based on values of two other variables

We will again continue using the `bechdel_00_13` data set you created two exercises ago.  It has been loaded for you here.

Let's now repeat the same analysis as before but look at the mean and median ROI values across both the `decade_code` and `clean_test` variables
(in that order).  To help with understanding of the results, `decade_code` is 1 for `2010-13` and 2 for `2000-2009`.

*** =instructions
- Determine the mean and median values for `roi_dom` across `decade_code` and `clean_test`.
- Call these new variables in the aggregated data frame `mean_roi` and `median_roi` respectively.
- Sort based on values of `mean_roi` with higher values at the top of the table and lower values at the bottom.
- Give the name of `roi_summ2` to the aggregated data frame and print it to the Console.

*** =hint
- You should be using `group_by` and `summarize` here.

*** =pre_exercise_code
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel <- bechdel %>%
    mutate(clean_test = factor(clean_test, 
                levels = c("nowomen", "notalk", "men", "dubious", "ok")))
bechdel_00_13 <- bechdel %>%
    filter(between(year, 2000, 2013)) %>%
    mutate(roi_dom = domgross_2013 / budget)
bechdel_00_13 <- bechdel_00_13 %>%
    mutate(roi_dom = budget_2013 / budget)    
```

*** =sample_code
```{r}

```

*** =solution
```{r}
library(dplyr)
roi_summ2 <- bechdel_00_13 %>%
    group_by(decade_code, clean_test) %>%
    summarize(mean_roi = mean(roi_dom),
              median_roi = median(roi_dom)) %>%
    arrange(desc(mean_roi))
roi_summ2
```

*** =sct
```{r}
test_library_function("dplyr")
test_object("roi_summ2")
success_msg("Nicely done! <br> <br> THINK ABOUT IT:  What stands out to you as you look at these numbers?  Which `clean_test` level
had the highest mean return on investment?  For 2010-13?  For 2000-2009?")
test_error()
```

--- type:NormalExercise lang:r xp:100 skills:1,4 key:ba670d6c5e
## Plot the results

We now shift to visualizing the data you created in the last exercise:  `roi_summ2`.  It has been loaded for you.

*** =instructions
- Write the code to produce the plot shown.

*** =hint
- 

*** =pre_exercise_code
```{r}
library(fivethirtyeight)
library(dplyr)
library(ggplot2)
bechdel <- bechdel %>%
    mutate(clean_test = factor(clean_test, 
                levels = c("nowomen", "notalk", "men", "dubious", "ok")))
bechdel_00_13 <- bechdel %>%
    filter(between(year, 2000, 2013)) %>%
    mutate(roi_dom = domgross_2013 / budget)
bechdel_00_13 <- bechdel_00_13 %>%
    mutate(roi_dom = budget_2013 / budget)    
roi_summ2 <- bechdel_00_13 %>%
    group_by(decade_code, clean_test) %>%
    summarize(mean_roi = mean(roi_dom),
              median_roi = median(roi_dom)) %>%
    arrange(desc(mean_roi))    
ggplot(data = roi_summ2, 
        mapping = aes(x = factor(decade_code), 
            y = mean_roi, fill = clean_test)) + 
    geom_col(position = "dodge", color = "white")
detach("package:ggplot2", unload = TRUE)  
```

*** =sample_code
```{r}

```

*** =solution
```{r}
library(ggplot2)
ggplot(data = roi_summ2, 
        mapping = aes(x = factor(decade_code), 
            y = mean_roi, fill = clean_test)) + 
    geom_col(position = "dodge", color = "white")
```

*** =sct
```{r}
test_library_function("ggplot2")
test_or({
  # match original solution
  ggplot_fun <- ex() %>% check_function("ggplot")
  ggplot_fun %>% check_arg("data") %>% check_equal(eval = FALSE)
  ggplot_fun %>% check_arg("mapping")
  aes_fun <- ex() %>% check_function("aes")
  aes_fun %>% check_arg("x") %>% check_equal(eval = FALSE)
  aes_fun %>% check_arg("y") %>% check_equal(eval = FALSE) 
  aes_fun %>% check_arg("fill") %>% check_equal(eval = FALSE)
  geom_bar_fun <- ex() %>% check_function("geom_col")
  geom_bar_fun %>% check_arg("position") %>% check_equal(eval = FALSE)
  geom_bar_fun %>% check_arg("color") %>% check_equal(eval = FALSE)
}, {
  # match solution 1:
  sol_code1 <- 'ggplot(data = roi_summ2, 
        mapping = aes(x = factor(decade_code), 
            y = mean_roi)) + 
    geom_col(mapping = aes(fill = clean_test), position = "dodge", color = "white")'
  alt_ex_1 <- ex() %>% override_solution(sol_code1)

  ggplot_fun <- alt_ex_1 %>% check_function("ggplot")
  ggplot_fun %>% check_arg("data") %>% check_equal(eval = FALSE)
  ggplot_fun %>% check_arg("mapping")
  aes_fun <- alt_ex_1 %>% check_function("aes")
  aes_fun %>% check_arg("x") %>% check_equal(eval = FALSE)
  aes_fun %>% check_arg("y") %>% check_equal(eval = FALSE)
  geom_bar_fun <- alt_ex_1 %>% check_function("geom_col")
  geom_bar_fun %>% check_arg("position") %>% check_equal(eval = FALSE)
  geom_bar_fun %>% check_arg("color") %>% check_equal(eval = FALSE)
  geom_bar_fun %>% check_arg("mapping")
  alt_ex_1 %>% check_function("aes", index = 2) %>% check_arg("fill") %>% check_equal(eval = FALSE)
})
success_msg("Great work!")
test_error()
```

--- type:NormalExercise lang:r xp:100 skills:1,4 key:04a65758c4
## What proportions were big?

Recall that you can calculate proportions using the `mean` function.  This works because when R sees something like `my_vec <= 10` it will produce a new vector of the same length as `my_vec` with values of either `TRUE` or `FALSE` corresponding to whether each entry in the vector is less than or equal to 10 or not.

Let's investigate what proportion of movies that passed and failed the Bechdel test in 2013 were big budget films with big international gross earnings.

*** =instructions
- Calculate the proportion of movies that had a 2013 budget of greater than 100 million (call it `prop_big_budg`) and the proportion of movies that had **international** gross earnings in 2013 or more than 400 million (call it `prop_big_gross`).  (Note that you can use scientific notation to write 100 million as `1e+08` and 400 million as `4e+08` in R if you'd rather not count zeroes.)
- Assign the name `bech_2013_big` to this aggregated data frame.  It should have two rows and three columns.
- Some of the entries for `intgross_2013` are missing so you should use `na.omit()` in your piping as well to remove all rows that have a column with a missing value as was done in the **Filtering, Grouping, & Summarizing** chapter.  If you don't, you'll get a value of `NA` for one of the entries of `prop_big_gross`.

*** =hint
- 

*** =pre_exercise_code
```{r}
library(fivethirtyeight)
library(dplyr)
bechdel <- bechdel %>%
    mutate(clean_test = factor(clean_test, 
                levels = c("nowomen", "notalk", "men", "dubious", "ok")))
```

*** =sample_code
```{r}
library(dplyr)

```

*** =solution
```{r}
library(dplyr)
bech_2013_big <- bechdel %>% 
    na.omit() %>%
    filter(year == 2013) %>% 
    group_by(binary) %>% 
    summarize(prop_big_budg = mean(budget_2013 > 1e+08), 
        prop_big_gross = mean(intgross_2013 > 4e+08))
```

*** =sct
```{r}
test_library_function("dplyr")
test_object("bech_2013_big")
success_msg("Great work! <br><br> THINK ABOUT IT:  What do these results tell you about how 2013 movies that passed the Bechdel test compared to those that didn't in terms of big budget, high grossing movies?")
test_error()
```


