---
title: "ModernDive data frames"
author: "Chester Ismay"
output: 
  html_document:
    toc: true
    toc_float: true
---

This page provides an interactive way to view the data frames used in the [ModernDive](http://moderndive.com) textbook.  It will take a few seconds to load initially.  

You can click on the **Chapter** the specific data frame is introduced in the floating Table of Contents.  The specific
package the data frame is included in is provided as a secondary header.  

These interactive data tables were created using the [`DT` R package](https://rstudio.github.io/DT/).  For data frames larger than 100 rows, the data has been split onto multiple
pages for ease of use. For data frames larger than 5000 rows, a random sample of size 5000 of the data
is selected instead.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE)
library(DT)
library(nycflights13)
library(dplyr)
library(mosaic)

interact_df <- function(data){
  
  if(nrow(data) > 100){
    datatable(resample(data, size = min(5000, nrow(data)), replace = FALSE),
      class = 'cell-border stripe',
      rownames = FALSE,
      filter = list(position = 'top'))
  } else {
  datatable(data,
    class = 'cell-border stripe',
    rownames = FALSE,
    filter = list(position = 'top'),
    options = list(pageLength = nrow(data), 
      autoWidth = TRUE,
      bInfo = FALSE,
      paging = FALSE))
  }
}
```



# Chapter 3

<p style="font-size:36px">`nycflights13`</p>

## `flights`

```{r}
interact_df(flights)
```

***

## `weather`

```{r}
interact_df(weather)
```

***

## `planes`

```{r}
interact_df(planes)
```

***

## `airports`

```{r}
interact_df(airports)
```

***

## `airlines`

```{r}
interact_df(airlines)
```

***

<p style="font-size:36px">Derived data</p>

## `joined_flights`

```{r}
joined_flights <- inner_join(x = flights, y = airlines, by = "carrier")
interact_df(joined_flights)
```

# Chapter 4

<p style="font-size:36px">Derived data</p>

## `all_alaska_flights`

```{r}
library(nycflights13)
all_alaska_flights <- flights %>% 
  filter(carrier == "AS")
interact_df(all_alaska_flights)
```

***

## `early_january_weather`

```{r}
early_january_weather <- weather %>% 
  filter(origin == "EWR" & month == 1 & day <= 15)
interact_df(early_january_weather)
```

***

## `flights_namedports`

```{r}
flights_namedports <- flights %>% 
  inner_join(airports, by = c("origin" = "faa"))
interact_df(flights_namedports)
```

